Unified treatment of the asymptotics of asymmetric kernel density estimators

### 2016-01-05 paper/paper.tex
Prepare publication for AOS.:

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass{amsart}
\usepackage[latin9]{inputenc}
\usepackage{rotfloat}
\usepackage{amstext}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage[backend=bibtex,style=numeric,sorting=none,firstinits=true]{biblatex}
\addbibresource{paper.bib}
\usepackage[foot]{amsaddr}

@@

\title[Unified treatment of the asymptotics of asymmetric KDEs]{Unified treatment of the asymptotics of asymmetric kernel density estimators}

\author{Till Hoffmann}

\address{Department of Mathematics, Imperial College London, SW7 2AZ, United Kingdom}

\email{t.hoffmann13@imperial.ac.uk}

\author{Nick Jones}

\email{n.jones@imperial.ac.uk}
\begin{abstract}
We extend balloon and sample-smoothing estimators, two types of variable-bandwidth kernel density estimators, by a shift parameter and derive their asymptotic properties. Our approach facilitates the unified study of a wide range of density estimators which are subsumed under these two general classes of kernel density estimators. We demonstrate our method by deriving the asymptotic bias, variance, and mean (integrated) squared error of density estimators with gamma, log-normal, Birnbaum-Saunders, inverse Gaussian and reciprocal inverse Gaussian kernels. We propose two new density estimators for positive random variables that yield properly-normalised density estimates. Plugin expressions for bandwidth estimation are provided to facilitate easy exploratory data analysis.
\end{abstract}

\maketitle

@@

\begin{sidewaystable}
\vspace{.65\textwidth}%
\begin{tabular*}{0.9\textheight}{@{\extracolsep{\fill}}llllll}
Distribution & $\left\langle t\right\rangle $ & Variance & Parametrisation & $h^{2}\left(x\right)\delta\left(x\right)$ & $h\left(x\right)$\tabularnewline
\hline 
$\gammakernel\left(t;k,\theta\right)=\frac{\theta^{-k}t^{k-1}}{\Gamma(k)}\exp\left(-\frac{t}{\theta}\right)$ & $k\theta$ & $k\theta^{2}$ & $\gammakernel\left(y;\frac{x}{\sigma^{2}}+1,\sigma^{2}\right)$ & $\sigma^{2}$ & $\sigma\sqrt{x+\sigma^{2}}$\tabularnewline
$\lognormalkernel\left(t;\mu,\sigma^{2}\right)=\frac{1}{\sqrt{2\pi}\sigma x}\exp\left(-\frac{\left(\log t-\mu\right)^{2}}{2\sigma^{2}}\right)$ & $\euler^{\mu+\frac{\sigma^{2}}{2}}$ & $\left\langle t\right\rangle ^{2}\left(\euler^{\sigma^{2}}-1\right)$ & $\lognormalkernel\left(y;\log x,\sigma^{2}\right)$ & $x\left(\euler^{\frac{\sigma^{2}}{2}}-1\right)$ & $x\euler^{\frac{\sigma^{2}}{2}}\sqrt{\euler^{\sigma^{2}}-1}$\tabularnewline
$\bskernel\left(t;\sigma,\lambda\right)=\frac{(1+t\lambda)}{2\sigma\sqrt{2\pi\lambda t^{3}}}\exp\left(-\frac{(t\lambda-1)^{2}}{2t\lambda\sigma^{2}}\right)$ & $\frac{2+\sigma^{2}}{2\lambda}$ & $\frac{\sigma^{2}\left(4+5\sigma^{2}\right)}{4\lambda^{2}}$ & $\bskernel\left(y;\sigma,x^{-1}\right)$ & $\frac{x\sigma^{2}}{2}$ & $x\sigma\sqrt{1+5\sigma^{2}/4}$\tabularnewline
$\igkernel\left(t;\mu,\lambda\right)=\sqrt{\frac{\lambda}{2\pi t^{3}}}\exp\left(-\frac{\lambda(t-\mu)^{2}}{2t\mu^{2}}\right)$ & $\mu$ & $\frac{\mu^{3}}{\lambda}$ & $\igkernel\left(y;x,\sigma^{-2}\right)$ & 0 & $\sigma x^{3/2}$\tabularnewline
$\rigkernel\left(t;\mu,\lambda\right)=\sqrt{\frac{\lambda}{2\pi t}}\exp\left(-\frac{\lambda(\mu t-1)^{2}}{2t\mu^{2}}\right)$ & $\frac{1}{\mu}+\frac{1}{\lambda}$ & $\frac{1}{\lambda\mu}+\frac{2}{\lambda^{2}}$ & $\rigkernel\left(y;\frac{1}{x-\sigma^{2}},\sigma^{-2}\right)$ & 0 & $\sigma\sqrt{x+\sigma^{2}}$\tabularnewline
\end{tabular*}

@@

\section{\label{app:proofs}Proofs}
\begin{lem}
\label{lem:integral-balloon}The integral

@@

\section{Cross-validation\label{app:cross-validation}}

The quantity 
\begin{equation}
M=\int dx\,\hat{f}^{2}\left(x\right)-\frac{2}{n}\sum_{i=1}^{n}\hat{f}_{i}\left(X_{i}\right)\label{eq:mise-estimator}
\end{equation}
is an unbiased estimator of the MISE of the density estimator $\hat{f}\left(x\right)$, where $\hat{f_{i}}\left(X_{i}\right)$ is the density estimated from all samples except $X_{i}$ evaluated at $X_{i}$. Details can be found in Section~3.4.3 of \cite{Silverman1986}. The second term in Eq.~\eqref{eq:mise-estimator} is straightforward to evaluate whereas the first requires integration of the density estimate over the whole domain. Fortunately, the integral is tractable for the proper gamma estimator in Eq.~\eqref{eq:jeon-kim-estimator}:
\begin{align*}
\int dx\,\hat{f}\left(x\right) & =\frac{1}{n^{2}}\sum_{i,j=1}^{n}\int dx\,\mathrm{G}\left(x;1+\frac{X_{i}}{\sigma^{2}},\sigma^{2}\right)\mathrm{G}\left(x;1+\frac{X_{j}}{\sigma^{2}},\sigma^{2}\right)\\
 & =\frac{1}{n^{2}}\sum_{i,j=1}^{n}\frac{\Gamma\left(1+\frac{X_{i}+X_{j}}{\sigma^{2}}\right)2^{-1-\frac{X_{i}+X_{j}}{\sigma^{2}}}}{\Gamma\left(1+X_{i}/\sigma^{2}\right)\Gamma\left(1+X_{j}/\sigma^{2}\right)}.
\end{align*}

\printbibliography
\end{document}

>>>>

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside,aos,preprint]{imsart}
\usepackage[latin9]{inputenc}
\usepackage{rotfloat}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{hyperref}

@@

\title{Unified treatment of the asymptotics of asymmetric kernel density estimators}
\author{Till Hoffmann\corref{}\ead[label=e1]{t.hoffmann13@imperial.ac.uk}}
\and
\author{Nick Jones\ead[label=e2]{nick.jones@imperial.ac.uk}}
\affiliation{Imperial College London}
\address{Department of Mathematics\\Imperial College London\\London, SW7 2AZ\\United Kingdom\\\printead{e1}\\ \phantom{E-mail:\ }\printead*{e2}}

\maketitle
\begin{abstract}
We extend balloon and sample-smoothing estimators, two types of variable-bandwidth kernel density estimators, by a shift parameter and derive their asymptotic properties. Our approach facilitates the unified study of a wide range of density estimators which are subsumed under these two general classes of kernel density estimators. We demonstrate our method by deriving the asymptotic bias, variance, and mean (integrated) squared error of density estimators with gamma, log-normal, Birnbaum-Saunders, inverse Gaussian and reciprocal inverse Gaussian kernels. We propose two new density estimators for positive random variables that yield properly-normalised density estimates. Plugin expressions for bandwidth estimation are provided to facilitate easy exploratory data analysis.
\end{abstract}

@@

\begin{sidewaystable}
\begin{tabular*}{0.9\textheight}{@{\extracolsep{\fill}}llllll}
Distribution & $\left\langle t\right\rangle $ & Variance & Parametrisation & $h^{2}\left(x\right)\delta\left(x\right)$ & $h\left(x\right)$\tabularnewline
\hline 
$\gammakernel\left(t;k,\theta\right)=\frac{\theta^{-k}t^{k-1}}{\Gamma(k)}\exp\left(-\frac{t}{\theta}\right)$ & $k\theta$ & $k\theta^{2}$ & $\gammakernel\left(y;\frac{x}{\sigma^{2}}+1,\sigma^{2}\right)$ & $\sigma^{2}$ & $\sigma\sqrt{x+\sigma^{2}}$\tabularnewline
$\lognormalkernel\left(t;\mu,\sigma^{2}\right)=\frac{1}{\sqrt{2\pi}\sigma x}\exp\left(-\frac{\left(\log t-\mu\right)^{2}}{2\sigma^{2}}\right)$ & $\euler^{\mu+\frac{\sigma^{2}}{2}}$ & $\left\langle t\right\rangle ^{2}\left(\euler^{\sigma^{2}}-1\right)$ & $\lognormalkernel\left(y;\log x,\sigma^{2}\right)$ & $x\left(\euler^{\frac{\sigma^{2}}{2}}-1\right)$ & $x\euler^{\frac{\sigma^{2}}{2}}\sqrt{\euler^{\sigma^{2}}-1}$\tabularnewline
$\bskernel\left(t;\sigma,\lambda\right)=\frac{(1+t\lambda)}{2\sigma\sqrt{2\pi\lambda t^{3}}}\exp\left(-\frac{(t\lambda-1)^{2}}{2t\lambda\sigma^{2}}\right)$ & $\frac{2+\sigma^{2}}{2\lambda}$ & $\frac{\sigma^{2}\left(4+5\sigma^{2}\right)}{4\lambda^{2}}$ & $\bskernel\left(y;\sigma,x^{-1}\right)$ & $\frac{x\sigma^{2}}{2}$ & $x\sigma\sqrt{1+5\sigma^{2}/4}$\tabularnewline
$\igkernel\left(t;\mu,\lambda\right)=\sqrt{\frac{\lambda}{2\pi t^{3}}}\exp\left(-\frac{\lambda(t-\mu)^{2}}{2t\mu^{2}}\right)$ & $\mu$ & $\frac{\mu^{3}}{\lambda}$ & $\igkernel\left(y;x,\sigma^{-2}\right)$ & 0 & $\sigma x^{3/2}$\tabularnewline
$\rigkernel\left(t;\mu,\lambda\right)=\sqrt{\frac{\lambda}{2\pi t}}\exp\left(-\frac{\lambda(\mu t-1)^{2}}{2t\mu^{2}}\right)$ & $\frac{1}{\mu}+\frac{1}{\lambda}$ & $\frac{1}{\lambda\mu}+\frac{2}{\lambda^{2}}$ & $\rigkernel\left(y;\frac{1}{x-\sigma^{2}},\sigma^{-2}\right)$ & 0 & $\sigma\sqrt{x+\sigma^{2}}$\tabularnewline
\end{tabular*}

@@

\section{Proofs}

\label{app:proofs}
\begin{lem}
\label{lem:integral-balloon}The integral

@@

\section{Cross-validation}

\label{app:cross-validation}The quantity 
\begin{equation}
M=\int dx\,\hat{f}^{2}\left(x\right)-\frac{2}{n}\sum_{i=1}^{n}\hat{f}_{i}\left(X_{i}\right)\label{eq:mise-estimator}
\end{equation}
is an unbiased estimator of the MISE of the density estimator $\hat{f}\left(x\right)$, where $\hat{f_{i}}\left(X_{i}\right)$ is the density estimated from all samples except $X_{i}$ evaluated at $X_{i}$. Details can be found in Section~3.4.3 of \cite{Silverman1986}. The second term in Eq.~\eqref{eq:mise-estimator} is straightforward to evaluate whereas the first requires integration of the density estimate over the whole domain. Fortunately, the integral is tractable for the proper gamma estimator in Eq.~\eqref{eq:jeon-kim-estimator}:
\begin{align*}
\int dx\,\hat{f}\left(x\right) & =\frac{1}{n^{2}}\sum_{i,j=1}^{n}\int dx\,\mathrm{G}\left(x;1+\frac{X_{i}}{\sigma^{2}},\sigma^{2}\right)\mathrm{G}\left(x;1+\frac{X_{j}}{\sigma^{2}},\sigma^{2}\right)\\
 & =\frac{1}{n^{2}}\sum_{i,j=1}^{n}\frac{\Gamma\left(1+\frac{X_{i}+X_{j}}{\sigma^{2}}\right)2^{-1-\frac{X_{i}+X_{j}}{\sigma^{2}}}}{\Gamma\left(1+X_{i}/\sigma^{2}\right)\Gamma\left(1+X_{j}/\sigma^{2}\right)}.
\end{align*}

\bibliographystyle{imsart-number}
\bibliography{paper}

\end{document}
